A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-17 shanghai.tex
Finished work on base change section

\begin{proposition} \label{prop:csbe}
Set $G_{k'} = G\times_\Spec{k} \Spec{k'}$ and let
\[
\QC(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota^*} \QC(G)
\]
be the functor obtained by pull-back along the canonical closed immersion of $k$-schemes  $\iota : G \hookrightarrow \Res_{k'/k}(G_{k'})$ \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6}. 
The following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota^*} \dar[swap]{\trFrob{\Res_{k'/k}(G_{k'})}} & \QCiso{G} \dar{\trFrob{G}} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
Suppose $\qcs{L} = (\gqcs{L}, \mu, \phi) \in \QC(\Res_{k'/k}(G_{k'}))$ and $x \in G(\Fq)$.
The value $t_G(\iota^*(\qcs{L}))(x)$ is given by the trace of $\iota^*\phi$ on $(\iota^*\gqcs{L})_x$.
Similarly, $\iota_{k'/k}^*(t_{\Res_{k'/k}(G_{k'})}(\qcs{L}))(x)$ is given by the trace of $\phi$ on
$\gqcs{L}_{\iota(x)}$.  These are the same since $\iota$ is an inclusion and $\phi$
preserves the stalk at $\iota(x)$. \David{Is this the right justification?}
\end{proof}

In the opposite direction, let $\Nm_{k'/k} : G(k') \to G(k)$ be the norm map and consider the group homomorphism:
\[
\Nm_{k'/k}^* : G(k)^* \to G(k')^* \qquad \text{defined by}\qquad \chi \mapsto \chi\circ \Nm_{k'/k}.
\]
We can also interpret this operation on characters in terms of quasicharacter sheaves, as follows.

If $\qcs{L} = (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$ we may define
a quasicharacter sheaf $\qcs{L}' = (\gqcs{L}, \mu, \phi_{k'})$ on 
$G_{k'}\ceq G\times_{\Spec{k}} \Spec{k'}$ by setting
\[
\phi_{k'} = \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (CS.2) for $\phi_{k'}$ 
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$ 
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
 The rule $\nu_{k'/k}: (\gqcs{L}, \mu, \phi) \mapsto (\gqcs{L}, \mu, \phi_{k'})$ 
 defines a monoidal functor $\QC(G) \to \QC(G_{k'})$. 
 Moreover, the following diagram commutes, 
 where the top map is induced by $\QC(G) \to \QC(G_{k'})$ 
 and the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}[column sep=60]
\QCiso{G} \rar{\nu_{k'/k}} \dar{\trFrob{G}} & \QCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm_{k'/k}^*} & G(k')^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
Let $\qcs{L} = (\gqcs{L}, \mu, \phi) \in \QC(G)$ and $x \in G(k')$.  The value $t_{G_{k'}}(\nu_{k'/k}(\qcs{L})(x)$ is given by the trace of $\phi_{k'}$ on $\gqcs{L}_x$, while the value $\Nm_{k'/k}^*(t_G(\qcs{L}))(x)$ is given by the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.  After choosing bases of $\gqcs{L}_x, \ldots, \gqcs{L}_{\Frob{G}^n(x)}$, an iteration of \ref{CS.2} applied to the stalk of $\gqcs{L}^{\boxtimes n}$ at $(x, \Frob{G}(x), \ldots, \Frob{G}^{n-1}(x))$ expresses the trace of $\phi$ at $\Nm_{k'/k}(x)$ as a product of 
\end{proof}

Finally, we explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
The component corresponding to the identity element yields a natural inclusion 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
of $k'$-schemes.  The definition of base change yields a map of $k$-schemes
\[
(\Res_{k'/k}G')_{k'} \to \Res_{k'/k}G'.
\]
Consecutive pullback along these maps yields a functor
\[
\QC(G') \to \QC(\Res_{k'/k}G').
\]

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\QCiso{G'} \dar{\trFrob{G'}} \rar & \QCiso{\Res_{k'/k} G'} \dar{\trFrob{\Res_{k'/k} G'}} \\
 G'(k')^* \rar & (\Res_{k'/k}G')(k)^*
\end{tikzcd}
\]
\end{proposition}

>>>>

\begin{proposition} \label{prop:csbe}
Set $G_{k'} = G\times_\Spec{k} \Spec{k'}$ and let
\[
\QC(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota^*} \QC(G)
\]
be the functor obtained by pull-back along the canonical closed immersion of $k$-schemes  $\iota : G \hookrightarrow \Res_{k'/k}(G_{k'})$ \cite{bosch-lutkebohmert-reynaud:NeronModels}*{\S 7.6}. 
The following diagram commutes:
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota^*} \dar[swap]{\trFrob{\Res_{k'/k}(G_{k'})}} & \QCiso{G} \dar{\trFrob{G}} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
This result follows immediately from Proposition~\ref{prop:pullback} together with the identifications
$\Res_{k'/k}(G_{k'})(k) \cong G_{k'}(k') \cong G(k')$ from the definitions of Weil restriction and base change.
\end{proof}

In the opposite direction, let $\Nm : G(k') \to G(k)$ be the norm map and consider the group homomorphism:
\[
\Nm^* : G(k)^* \to G(k')^* \qquad \text{defined by}\qquad \chi \mapsto \chi\circ \Nm.
\]
We can also interpret this operation on characters in terms of quasicharacter sheaves, as follows.

If $\qcs{L} = (\gqcs{L}, \mu, \phi)$ is a quasicharacter sheaf on $G$ we may define
a quasicharacter sheaf $\qcs{L}' = (\gqcs{L}, \mu, \phi_{k'})$ on the base change
$G_{k'}$ of $G$ to $k'$ by setting
\[
\phi_{k'} = \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (CS.2) for $\phi_{k'}$ 
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gqcs{L}$ 
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
 The rule $\nu_{k'/k}: (\gqcs{L}, \mu, \phi) \mapsto (\gqcs{L}, \mu, \phi_{k'})$ 
 defines a monoidal functor $\QC(G) \to \QC(G_{k'})$. 
 Moreover, the following diagram commutes, 
 where the top map is induced by $\QC(G) \to \QC(G_{k'})$ 
 and the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}[column sep=60]
\QCiso{G} \rar{\nu_{k'/k}} \dar{\trFrob{G}} & \QCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm^*} & G(k')^*.
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
Let $\qcs{L} = (\gqcs{L}, \mu, \phi) \in \QC(G)$ and $x \in G(k')$ and write $\Frob{}$ for $\Frob{G}$.
The value $t_{G_{k'}}(\nu_{k'/k}(\qcs{L})(x)$ is given by the trace of $\phi_{k'}$ on $\gqcs{L}_x$,
while the value $\Nm^*(t_G(\qcs{L}))(x)$ is given by the trace of $\phi$ on $\gqcs{L}_{\Nm(x)}$.
Applying \ref{CS.2} to the stalk of $\gqcs{L}^{\boxtimes n}$ at $(x, \Frob{}(x), \ldots, \Frob{}^{n-1}(x))$ yields a diagram
\[
\begin{tikzcd}
\gqcs{L}_{\Nm(x)} \rar \dar{\phi_{\Nm(x)}} & \gqcs{L}_{\Frob{}(x)} \otimes \gqcs{L}_{\Frob{}^2(x)} \otimes \cdots \otimes \gqcs{L}_x
\dar{\phi_x \otimes (\Frob{}^*\phi)_x \otimes \cdots \otimes ((\Frob{}^{n-1})^*\phi)_x} \\
\gqcs{L}_{\Nm(x)} \rar & \gqcs{L}_x \otimes \gqcs{L}_{\Frob{}(x)} \otimes \cdots \otimes \gqcs{L}_{\Frob{}^{n-1}(x)}.
\end{tikzcd}
\]
Choose a basis vector $v$ for $\gqcs{L}_{\Nm(x)}$, mapping to $v_0 \otimes v_1 \otimes \cdots \otimes v_{n-1}$
along the bottom map, where $v_i \in \gqcs{L}_{\Frob{}^i(x)}$.  By \ref{CS.1}, $v$ maps to
$v_1 \otimes v_2 \otimes \cdots \otimes v_0$ along the top of the diagram.
Let $\alpha_i \in \EEx$ represent $((\Frob{}^i)^*\phi)_x$ with respect to these bases and let $\alpha$ be
the trace of $\phi_{\Nm(x)}$.  We may now equate the trace $\alpha$ of $\phi$ on $\gqcs{L}_{\Nm(x)}$
with the product $\alpha_0 \cdots \alpha_{n-1}$, which is the trace of $\phi_{k'}$ on $\gqcs{L}_x$.
\end{proof}

Finally, let $G'$ be a smooth commutative group scheme locally of finite type over $k'$;
we explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
Since products and coproducts agree for group schemes we have a natural inclusion of $k'$-schemes 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
mapping $G'$ into the summand corresponding to $1 \in \Gal(k'/k)$.  Composing pullback
along this map with $\nu_{k'/k}$ from Proposition~\ref{prop:basechange} yields a functor
\[
\QC(\Res_{k'/k}G') \to \QC(G').
\]

\begin{proposition}
The following diagram of isomorphisms commutes, where the bottom map is the identity:
\[
\begin{tikzcd}
\QCiso{\Res_{k'/k} G'} \dar{\trFrob{\Res_{k'/k} G'}} \rar & \QCiso{G'} \dar{\trFrob{G'}}\\
G'(k')^* \rar & G'(k')^*
\end{tikzcd}
\]
\end{proposition}
\begin{proof}
By Proposition~\ref{prop:pullback} the pullback part of the definition corresponds to the map $(\Res_{k'/k}G')(k') \to G'(k')^*$
induced by $g \mapsto (g, 1, \ldots, 1)$.  Since the action of $\Gal(k'/k)$ on $(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'$
is given by permuting coordinates, composition with the norm map yields the identity on $G'(k')$.
\end{proof}
